The Astropy Project: \\ Sustaining and Growing a Community-oriented Open-source Project and the Latest Major Release (v5.0) of the Core Package

### 2022-04-06 src/ms.tex

The APE0 \citep{ape0} document lays out the principles of this governance
structure, so we refer the reader to that document for a more thorough
description, and here we only highlight some key elements. While many of these
principles were already de facto true or have been discussed organically (and
have been discussed in earlier papers in this series), the APE0-based governance
aims to provide a single place where the community can agree as a starting
point. With this in mind, it highlights the developer and user community of the
Astropy Project as the ultimate sources of authority, as well as the core
principle of ``do-ocracy'': those who do work for the Project (be it coding,
training, or other harder-to-quantify contributions) gain more influence on the
outputs of the Project by virtue of their effort. However, APE0 adds the
concept of ``voting members'': a self-governed part of that community who are
entrusted to elect the Coordination Committee (CoCo). While the CoCo has existed
from the inception of the project, APE0 establishes a formal voting process for
CoCo members, and explicitly outlines the rights and responsibilities of the
CoCo. This role is mainly to facilitate consensus and act as the decision maker
when other mechanisms have failed, and includes powers that either
require central authority or secrets (e.g., passwords). However, APE0 also
charges the Committee to devolve responsibilities and seek community input on
these items as often as possible.

>>>>

The APE 0 \citep{ape0} document lays out the principles of this governance
structure, so we refer the reader to that document for a more thorough
description, and here we only highlight some key elements. While many of these
principles were already de facto true or have been discussed organically (and
have been discussed in earlier papers in this series), the APE 0-based governance
aims to provide a single place where the community can agree as a starting
point. With this in mind, it highlights the developer and user community of the
Astropy Project as the ultimate sources of authority, as well as the core
principle of ``do-ocracy'': those who do work for the Project (be it coding,
training, or other harder-to-quantify contributions) gain more influence on the
outputs of the Project by virtue of their effort. However, APE 0 adds the
concept of ``voting members'': a self-governed part of that community who are
entrusted to elect the Coordination Committee (CoCo). While the CoCo has existed
from the inception of the project, APE 0 establishes a formal voting process for
CoCo members, and explicitly outlines the rights and responsibilities of the
CoCo. This role is mainly to facilitate consensus and act as the decision maker
when other mechanisms have failed, and includes powers that either
require central authority or secrets (e.g., passwords). However, APE 0 also
charges the Committee to devolve responsibilities and seek community input on
these items as often as possible.
